\vspace{-\baselineskip} Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-2\baselineskip}

### 2017-05-31 paper/ms.tex

To generate a draw from a Gaussian process with a specified \celerite\ model requires carrying out the
low-rank Cholesky decomposition first, generating the vector ${\bf g}$ of random normal deviates, and then multiplying
\begin{equation}
{\bf y} = L D^{1/2} {\bf g}.
\end{equation}
This may be carried out in $O(JN)$ operations with
\begin{eqnarray}
f_{n,j} &=& \phi_{n-1,j} \left(f_{n-1,j} + \tilde{X}_{n-1,j} g_{n-1}\right)\\
y_n &=& D_{n,n}^{1/2} g_n + \sum_{k=1}^{2J} \tilde{U}_{n,k} f_{n,k}
\end{eqnarray}
for $n = 1$ to $N$ and $j=1$ to $2J$, with $f_{0,j} = 0$ for all $j$ and $g_0 = 0$.
We expect that this simulation algorithm will be useful for testing Gaussian
process inference algorithms with large datasets, which becomes prohibitive
for large $N$ if standard Cholesky decomposition and multiplication are used.

@@

>>>>

To generate a draw from a Gaussian process with a specified \celerite\ model requires carrying out the
low-rank Cholesky decomposition first, generating the vector ${\bf g}$ of random normal deviates, and then multiplying
\begin{equation}
{\bf y} = L D^{1/2} {\bf g}.
\end{equation}
This may be carried out in $O(JN)$ operations with
\begin{eqnarray}
f_{n,j} &=& \phi_{n-1,j} \left(f_{n-1,j} + \tilde{X}_{n-1,j} g_{n-1}\right)\\
y_n &=& D_{n,n}^{1/2} g_n + \sum_{k=1}^{2J} \tilde{U}_{n,k} f_{n,k}
\end{eqnarray}
for $n = 1$ to $N$ and $j=1$ to $2J$, with $f_{0,j} = 0$ for all $j$ and $g_0 = 0$.
\agoltodo{Check that these equations are in fact correct.}
We expect that this simulation algorithm will be useful for testing Gaussian
process inference algorithms with large datasets, which becomes prohibitive
for large $N$ if standard Cholesky decomposition and multiplication are used.

Multiplication of the full covariance matrix times a vector can be accomplished
by yet another multiplication by the transpose of the Cholesky decomposition.
\dfmtodo{Should we spell this out as well?}

@@

\agoltodo{Write up the prediction algorithm.}
